---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First we need to unzip then load the data into a dataframe. Since we are working from a github repository it is safe to assume the zip file exists.
```{r load}
library(lubridate)
unzip("activity.zip")
activity <- read.csv("activity.csv", header = TRUE)
activity$date <- ymd(activity$date)
```

## What is mean total number of steps taken per day?
Now let's look at the skewness of our step counts.
```{r mean_steps}
library(ggplot2)
qplot(x=date,weight=steps,ylab="total steps", data=activity, geom="histogram")

mean_steps <- mean(activity$steps, na.rm=TRUE)
median_steps <- median(activity$steps, na.rm=TRUE)
```
 The take an average of `r mean_steps` steps daily, and their median step count is `r median_steps`.

## What is the average daily activity pattern?
```{r pattern}
library(dplyr)
pattern <- activity %>% group_by(interval) %>% summarise(average_steps=mean(steps,na.rm=TRUE))
ggplot(data=pattern,aes(x=interval,y=average_steps)) + geom_line()
max <- pattern %>% select(interval,average_steps) %>%  top_n(1,average_steps)

```

Our subjects are typically most active around interval `r max$interval` with an average of step count of `r max$average_steps`.
## Imputing missing values
Perhaps the missing records is affecting our results. Let's replace our NA values with the total average steps and compare with our previous results.
```{r imput}
sum(is.na(activity$steps))
activity2 <- activity
activity2$steps[is.na(activity2$steps)]<-mean(activity2$steps, na.rm=TRUE)

qplot(date,weight=steps, data=activity2, geom='histogram')
mean_steps2 <-mean(activity2$steps)
median_steps2<-median(activity2$steps)
```
After our adjustments our subjects took `r mean_steps2` steps daily, with a median of `r median_steps2`.

## Are there differences in activity patterns between weekdays and weekends?
To answer this question we need to first separate our weedays and weekends. We'll define and leverage a function called daysOrEnd to accomplish this.
```{r weekends}
library(lubridate)
dayOrEnd <- function(date){
  day_number <- wday(ymd(date))
  if(day_number == 1 | day_number == 7){
    day_type <- "weekend"
    }    
  else{
    day_type <- "weekday"
  } 
  day_type
  }

activity2$day_type <- as.factor(sapply(activity$date,dayOrEnd))
```
Now let's look at the activity patterns separated by weekdays and weekends.
```{r busy_weekend}
pattern <- activity2 %>% group_by(day_type,interval) %>% summarise(avg_steps=mean(steps))
ggplot(pattern,aes(x=interval,y=avg_steps))+geom_line()+facet_grid(day_type~.)
```

It looks on the weekends our subjects tend spread their steps out more.